---
title: "Ratings"
author: "Doruk Kilitcioglu"
date: "March 23, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Amazon vs Goodreads Ratings

We try to visualize the ratings from our two sources of data. Both of the csv files are in the standard user-item-rating format.

```{r}
library(tidyverse)
goodreads <- read_csv("../data/goodbooks-10k/ratings.csv")
head(goodreads)
```

```{r}
amazon <- read_csv("../data/amazon/ratings_amazon.csv")
head(amazon)
```

We first look at the difference in the distributions of the data.

```{r}
library(ggthemes)
all_ratings <- bind_rows(
  goodreads %>%
    mutate(Source = "Goodreads"),
  amazon %>%
    mutate(Source = "Amazon")
)
gsum <- goodreads %>%
  group_by(rating) %>%
  summarize(Count = n()) %>%
  mutate(Source = "Goodreads", Percent = Count/sum(Count)) %>%
  ungroup()
asum <- amazon %>%
  group_by(rating) %>%
  summarize(Count = n()) %>%
  mutate(Source = "Amazon", Percent = Count/sum(Count)) %>%
  ungroup()
csum <- all_ratings %>%
  group_by(rating, Source) %>%
  summarize(Count = n()) %>%
  group_by(Source) %>%
  mutate(Percent = Count/sum(Count))
head(csum)
```

```{r}
ggplot(csum, aes(x = reorder(rating, -rating), y = Percent, fill = Source)) +
  geom_bar(data = filter(csum, Source == "Goodreads"), stat = "identity") +
  geom_bar(data = filter(csum, Source == "Amazon") %>% mutate(Percent = -Percent), stat = "identity") +
  geom_text(data = filter(csum, Source == "Goodreads"), aes(label = sprintf("%.0f%%", Percent * 100), x = reorder(rating, -rating), y = Percent + 0.05)) +
  geom_text(data = filter(csum, Source == "Amazon"), aes(label = sprintf("%.0f%%", Percent * 100), x = reorder(rating, -rating), y = -Percent - 0.05)) +
  coord_flip() +
  scale_y_continuous(limits = c(-.7, .7), labels = NULL) +
  ylab("# Ratings") +
  xlab("Rating") +
  theme_fivethirtyeight() +
  theme(legend.title=element_blank()) +
  ggtitle("Relative distribution of ratings")
```

It is apparent that Amazon ratings are heavily biased towards 5 stars, while the Goodreads ratings are more balanced.

```{r}
all_ratings %>%
  group_by(user_id, Source) %>%
  summarize(Avg_Rating = mean(rating)) %>%
  group_by(Source) %>%
  summarize(Avg = mean(Avg_Rating))
```

The average ratings per user are also quite high. Even though the distribution looks very skewed for Amazon, the average ratings per user are not too different.

We also look at the distributions of user ratings.

```{r}
user_averages <- all_ratings %>%
  group_by(user_id, Source) %>%
  summarize(Avg_Rating = mean(rating), Num_Rated = n(), Variance = var(rating))

ggplot(user_averages, aes(x = Avg_Rating, fill = Source, color = Source)) +
  geom_density(alpha = .3, kernel = "gaussian", adjust = 1.8) +
  xlab("Average Rating per User") +
  ylab("") +
  theme_fivethirtyeight() +
  theme(legend.title=element_blank(),
        axis.text.y=element_blank()) +
  ggtitle("Average user rating distribution")
```

The Goodreads ratings look like a standard Gaussian curve, while the Amazon ratings look a bit more bimodal (see the little part sticking out to the right) and is much more concentrated (has less variance).

I also wanted to explore the correlation between the number of ratings of a user and the user's average rating.

```{r}
ggplot(filter(user_averages, Num_Rated <=200), aes(x = Num_Rated, y = Avg_Rating)) +
  geom_hex(bins = 60) +
  scale_fill_distiller(palette = "Spectral") +
  facet_grid(.~Source) +
  ylab("Average Rating per User") +
  xlab("Number of Ratings per User") +
  theme_minimal() +
  theme(strip.text = element_text(size = 14, face = "bold")) +
  labs(fill = "# Users")
```

As you can see, there doesn't seem to be much of a correlation, although we can clearly see how most users are centered around the average rating of 4 and around 100 books.

These results warn us against naively adding in both of the ratings together. At the very least, we need to normalize the ratings across the two datasets, but even that will not help with the fact that the distribution of the ratings are so different. It may even be in our best interests to not add the Amazon ratings, but we can only try and see.